---
title: "Somerville Happiness Survey – Final Exam Appendix"
output: html_notebook
---

This R Notebook provides the data preparation, model fitting, and diagnostic output for the Somerville Happiness Survey analyses described in the written report. 

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

##### Analysis 1: (Multiple Linear Regression)
```{r}
happiness <- read.csv("/Users/beatrizmorales/Downloads/happinessSurvey.csv", stringsAsFactors = TRUE)
head(happiness)
str(happiness)
summary(happiness)
```

```{r}
happiness$overall_happiness    <- happiness$HappyNow
happiness$life_satisfaction    <- happiness$GeneralLifeSatisfaction
names(happiness)
```

```{r}
happiness$neighborhood_quality <- happiness$NbrHoodSatisfaction
happiness$length_residence     <- happiness$YearsResidence
reg_data <- subset(
  happiness,
  !is.na(overall_happiness) &
  !is.na(neighborhood_quality) &
  !is.na(length_residence)
)
nrow(reg_data)
summary(reg_data[, c("overall_happiness",
                      "neighborhood_quality",
                      "length_residence")])
```

```{r}
model1 <- lm(overall_happiness ~ neighborhood_quality + length_residence,
             data = reg_data)

summary(model1)
```

```{r}
par(mfrow = c(2, 2))
plot(model1)
par(mfrow = c(1, 1))
```

#### Analysis 2: (Logistic Regression-Classification)

```{r}
happiness$high_life_satisfaction <- ifelse(happiness$life_satisfaction >= 8, 1, 0)
table(happiness$high_life_satisfaction)
```

```{r}
logit_data <- subset(
  happiness,
  !is.na(high_life_satisfaction) &
  !is.na(SafeWalkingatNight) &
  !is.na(neighborhood_quality)
)

nrow(logit_data)

summary(logit_data[, c("high_life_satisfaction",
                        "SafeWalkingatNight",
                        "neighborhood_quality")])
```

```{r}
model2 <- glm(high_life_satisfaction ~ SafeWalkingatNight + neighborhood_quality,
              data = logit_data,
              family = binomial)

summary(model2)
```

```{r}
logit_data$predicted_prob <- predict(model2, type = "response")
logit_data$predicted_class <- ifelse(logit_data$predicted_prob >= 0.5, 1, 0)

conf_matrix <- table(
  Actual = logit_data$high_life_satisfaction,
  Predicted = logit_data$predicted_class
)
conf_matrix
conf_matrix_df <- as.data.frame(conf_matrix)

library(ggplot2)

ggplot(conf_matrix_df, aes(x = Predicted, y = Actual, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), color = "white", size = 5) +
  labs(title = "Confusion Matrix Heatmap",
       x = "Predicted Class",
       y = "Actual Class") +
  theme_minimal()
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
accuracy
```

```{r}
library(pROC)

roc_obj <- roc(logit_data$high_life_satisfaction,
               logit_data$predicted_prob)

plot(roc_obj)
auc(roc_obj)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

