---
title: "Community Happiness & Life Satisfaction Analysis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(ggplot2)
library(pROC)

happiness <- read.csv("data/happinessSurvey.csv", stringsAsFactors=TRUE)
happiness$overall_happiness <- happiness$HappyNow
happiness$life_satisfaction <- happiness$GeneralLifeSatisfaction
happiness$neighborhood_quality <- happiness$NbrHoodSatisfaction
happiness$length_residence <- happiness$YearsResidence
```

## Multiple linear regression
```{r}
reg_data <- subset(
  happiness,
  !is.na(overall_happiness) &
  !is.na(neighborhood_quality) &
  !is.na(length_residence)
)

model1 <- lm(
  overall_happiness ~ neighborhood_quality + length_residence,
  data=reg_data
)
summary(model1)
par(mfrow=c(2,2)); plot(model1); par(mfrow=c(1,1))
```

## Logistic classification
The original coursework defined high life satisfaction as a score of 8 or higher.

```{r}
happiness$high_life_satisfaction <- ifelse(happiness$life_satisfaction >= 8, 1, 0)
logit_data <- subset(
  happiness,
  !is.na(high_life_satisfaction) &
  !is.na(SafeWalkingatNight) &
  !is.na(neighborhood_quality)
)

set.seed(1234)
idx <- sample(seq_len(nrow(logit_data)), floor(0.70*nrow(logit_data)))
train <- logit_data[idx, ]
test  <- logit_data[-idx, ]

model2 <- glm(
  high_life_satisfaction ~ SafeWalkingatNight + neighborhood_quality,
  data=train, family=binomial
)
summary(model2)
```

## Held-out evaluation
```{r}
test$predicted_prob <- predict(model2, newdata=test, type="response")
test$predicted_class <- ifelse(test$predicted_prob >= 0.5, 1, 0)

cm <- table(Actual=test$high_life_satisfaction, Predicted=test$predicted_class)
cm

tn <- cm["0","0"]; fp <- cm["0","1"]
fn <- cm["1","0"]; tp <- cm["1","1"]

metrics <- data.frame(
  Accuracy=(tp+tn)/sum(cm),
  Recall=tp/(tp+fn),
  Specificity=tn/(tn+fp),
  Precision=tp/(tp+fp)
)
metrics

roc_obj <- roc(test$high_life_satisfaction, test$predicted_prob, quiet=TRUE)
auc(roc_obj)
plot(roc_obj)
```

## Interpretation
Held-out evaluation is preferable to reporting classification performance on the same observations used to fit the model. AUC summarizes ranking ability across thresholds, while threshold-specific metrics show the consequences of the selected classification cutoff.
